KidsAIR wood stove grades and usage

### 2020-10-19 main_analysis_initial.Rmd
analysis updates

```{r, eval=TRUE, include=TRUE}
# Load individual datasets

@@

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  group_by(area, child_id_num) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
                                             person_time_no_lrti_total),
         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  ungroup() %>% 
  filter(person_time_at_risk_total > 0) %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence)
```

```{r}
# Merge dataset with blinded treatments and final LRTI data, select analysis variables
analysis_data <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  mutate(child_age = fct_collapse(age_child_nov,
                                  "< 1" = "Less than 1",
                                  "1+" = c(1, 2, 3, 4))) %>% 
  left_join(treatments_masked, by = "home") %>% 
  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
         person_time_at_risk_total, lrti_incidence, newtx) %>% 
  rename(treatment = newtx,
         lrti_events = lrti_events_overall,
         person_time_at_risk = person_time_at_risk_total) %>% 
  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
  filter(lrti_incidence < 0.7) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Dearborn", "Smith", "Blackfoot")))
```

@@

char_funct_total("gender_parent")
char_funct_treatment("gender_parent")
char_funct_total("hispanic_parent")
char_funct_treatment("hispanic_parent")
char_funct_total("race_parent")
char_funct_treatment("race_parent")
char_funct_total("education_3level")
char_funct_treatment("education_3level")
char_funct_total("income_3level")
char_funct_treatment("income_3level")
char_funct_total("home_floors_2level")
char_funct_treatment("home_floors_2level")
char_funct_total("home_year_built_2level")
char_funct_treatment("home_year_built_2level")
char_funct_total("home_pets_2level")
char_funct_treatment("home_pets_2level")
char_funct_total("home_pets")
char_funct_treatment("home_pets")
char_funct_total("stove_age_3level")
char_funct_treatment("stove_age_3level")
char_funct_total("chimney_clean_3level")
char_funct_treatment("chimney_clean_3level")
char_funct_total("wood_collect_method_2level")
char_funct_treatment("wood_collect_method_2level")
char_funct_total("wood_collect_2level")
char_funct_treatment("wood_collect_2level")
char_funct_total("stove_grade_3level")
char_funct_treatment("stove_grade_3level")
char_funct_total("burn_level_3level")
char_funct_treatment("burn_level_3level")
char_funct_total("ruca_code")
char_funct_treatment("ruca_code")
char_funct_total("ruca_code_2")
char_funct_treatment("ruca_code_2")
char_funct_total("ruca_code_3")
char_funct_treatment("ruca_code_3")
```

@@

num_funct_total("residents_under_five")
num_funct_treatment("residents_under_five")
num_funct_total("home_bedrooms")
num_funct_treatment("home_bedrooms")
num_funct_total("pm_mean_sampling_period")
num_funct_treatment("pm_mean_sampling_period")
num_funct_total("pm_sample_interval")
num_funct_treatment("pm_sample_interval")
num_funct_total("moisture_closest")
num_funct_treatment("moisture_closest")
num_funct_total("temp_indoor_max")
num_funct_treatment("temp_indoor_max")
num_funct_total("rh_indoor_max")
num_funct_treatment("rh_indoor_max")
num_funct_total("mean_temp")
num_funct_treatment("mean_temp")
num_funct_total("amb_pm_24hr")
num_funct_treatment("amb_pm_24hr")
num_funct_total("sums_events_sampling_period_2.5")
num_funct_treatment("sums_events_sampling_period_2.5")
num_funct_total("sums_events_sampling_period_5")
num_funct_treatment("sums_events_sampling_period_5")
num_funct_total("sums_events_sampling_period_7.5")
num_funct_treatment("sums_events_sampling_period_7.5")
```

@@

@@

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
    facet_wrap(~treatment)
events_di_plot

# Check for overdispersion in data using likelihood ratio test
library(pscl)
library(MASS)

nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), data = analysis_data)
odTest(nb_results)

@@

```

>>>>

```{r, eval=TRUE, include=TRUE}
# Load individual datasets, add new vars, and merge/select analysis dataset

@@

kids_amb_pm_data <- 
  read_rds(paste0(file_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = ymd(amb_pm_date))

kids_amb_temp_data <- 
  read_rds(paste0(file_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = ymd(temp_date))

lrti_final_data <- read_rds(paste0(file_path, "Output/lrti_final_data.rds")) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  mutate(cohort = round_date(csq_date, unit = "year")) %>% 
  separate(cohort, c("cohort", "trash"), sep = "-") %>% 
  group_by(area, child_id_num, winter_id) %>% 
  mutate(lrti_events = n()) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  group_by(area, child_id_num) %>% 
  mutate(lrti_events = as.numeric(lrti_events),
         lrti_events = if_else(is.na(person_time_no_lrti_total), 0, lrti_events),
         lrti_events_overall = sum(lrti_events, na.rm = TRUE),
         person_time_no_lrti_total = as.numeric(person_time_no_lrti_total),
         person_time_no_lrti_total = if_else(is.na(person_time_no_lrti_total), person_time_possible,
                                             person_time_no_lrti_total),
         person_time_at_risk_total = sum(person_time_no_lrti_total, na.rm = TRUE),
         person_time_at_risk_total = as.numeric(person_time_at_risk_total)/7,
         lrti_incidence = lrti_events_overall/person_time_at_risk_total,
         lrti_incidence = if_else(is.na(lrti_incidence), 0, lrti_incidence)) %>% 
  arrange(area, home, child_id_num, winter_id) %>% 
  left_join(kids_amb_pm_data, by = c("area", "home")) %>% 
  mutate(pm_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(pm_filter == 1) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_pm_lrti = mean(amb_pm_24hr, na.rm = TRUE)) %>% 
  distinct(area, child_id_num, winter_id, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, person_time_start, person_time_end) %>% 
  left_join(kids_amb_temp_data, by = c("area", "home")) %>% 
  mutate(temp_filter = if_else(sample_date >= person_time_start & sample_date <= person_time_end, 1, 0)) %>% 
  filter(temp_filter == 1 | child_id_num == 94) %>% 
  group_by(area, child_id_num) %>% 
  mutate(amb_temp_lrti = mean(mean_temp, na.rm = TRUE),
         amb_temp_lrti = if_else(child_id_num == 94, NaN, amb_temp_lrti),
         amb_dewp_lrti = mean(mean_dewp, na.rm = TRUE),
         amb_dewp_lrti = if_else(child_id_num == 94, NaN, amb_dewp_lrti),
         amb_pressure_lrti = mean(mean_pressure, na.rm = TRUE),
         amb_pressure_lrti = if_else(child_id_num == 94, NaN, amb_pressure_lrti),
         amb_wind_lrti = mean(mean_wind_knots, na.rm = TRUE),
         amb_wind_lrti = if_else(child_id_num == 94, NaN, amb_wind_lrti)) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  ungroup() %>% 
  select(area, home, cohort, child_id_num, lrti_events_overall, person_time_at_risk_total, lrti_incidence,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti)

# Merge dataset with blinded treatments and final LRTI data, select analysis variables
kids_analysis_data_1obs_per_child <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  mutate(child_age = fct_collapse(age_child_nov,
                                  "< 1" = "Less than 1",
                                  "1+" = c(1, 2, 3, 4))) %>% 
  left_join(treatments_masked, by = "home") %>% 
  arrange(area, home, child_id_num, winter_id, age_child_nov) %>% 
  right_join(lrti_final_data, by = c("area", "home", "child_id_num")) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  select(area, home, cohort, child_id_num, child_age, lrti_events_overall, 
         person_time_at_risk_total, lrti_incidence, newtx,
         amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti,
         gender_parent, hispanic_parent, race_parent, education_3level, income_3level,
         home_floors_2level, home_year_built_2level, home_pets_2level, home_pets,
         stove_age_3level, chimney_clean_3level, wood_collect_method_2level,
         wood_collect_2level, stove_grade_3level, burn_level_3level, ruca_code,
         ruca_code_2, ruca_code_3, residents_smoke, residents_smoke_inside,
         residents_under_five, home_bedrooms, pm_mean_sampling_period, pm_sample_interval,
         moisture_closest, temp_indoor_max, rh_indoor_max, sums_events_sampling_period_2.5,
         sums_events_sampling_period_5, sums_events_sampling_period_7.5) %>% 
  rename(treatment = newtx,
         lrti_events = lrti_events_overall,
         person_time_at_risk = person_time_at_risk_total) %>% 
  mutate(log_person_time_at_risk = log(person_time_at_risk)) %>% 
  mutate(lrti_events_di = if_else(lrti_events == 0, 0, 1)) %>% 
  mutate(treatment = factor(treatment,
                            levels = c("Blackfoot", "Dearborn", "Smith")))

write_rds(kids_analysis_data_1obs_per_child, 
          paste0(file_path, "Output/kids_analysis_data_1obs_per_child.rds"))
write_csv(kids_analysis_data_1obs_per_child, 
          paste0(file_path, "Output/kids_analysis_data_1obs_per_child.csv"), na = "")
```

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

analysis_data <- read_rds(paste0(file_path, 
                                 "Output/kids_analysis_data_1obs_per_child.rds"))

treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))
```

@@

char_funct_total("gender_parent")
char_funct_treatment("gender_parent")
char_funct_total("hispanic_parent")
char_funct_treatment("hispanic_parent")
char_funct_total("race_parent")
char_funct_treatment("race_parent")
char_funct_total("education_3level")
char_funct_treatment("education_3level")
char_funct_total("income_3level")
char_funct_treatment("income_3level")
char_funct_total("home_floors_2level")
char_funct_treatment("home_floors_2level")
char_funct_total("home_year_built_2level")
char_funct_treatment("home_year_built_2level")
char_funct_total("home_pets_2level")
char_funct_treatment("home_pets_2level")
char_funct_total("home_pets")
char_funct_treatment("home_pets")
char_funct_total("stove_age_3level")
char_funct_treatment("stove_age_3level")
char_funct_total("chimney_clean_3level")
char_funct_treatment("chimney_clean_3level")
char_funct_total("wood_collect_method_2level")
char_funct_treatment("wood_collect_method_2level")
char_funct_total("wood_collect_2level")
char_funct_treatment("wood_collect_2level")
char_funct_total("stove_grade_3level")
char_funct_treatment("stove_grade_3level")
char_funct_total("burn_level_3level")
char_funct_treatment("burn_level_3level")
char_funct_total("ruca_code")
char_funct_treatment("ruca_code")
char_funct_total("ruca_code_2")
char_funct_treatment("ruca_code_2")
char_funct_total("ruca_code_3")
char_funct_treatment("ruca_code_3")
char_funct_total("residents_smoke")
char_funct_treatment("residents_smoke")
char_funct_total("residents_smoke_inside")
char_funct_treatment("residents_smoke_inside")
```

@@

num_funct_total("residents_under_five")
num_funct_treatment("residents_under_five")
num_funct_total("home_bedrooms")
num_funct_treatment("home_bedrooms")
num_funct_total("pm_mean_sampling_period")
num_funct_treatment("pm_mean_sampling_period")
num_funct_total("pm_sample_interval")
num_funct_treatment("pm_sample_interval")
num_funct_total("moisture_closest")
num_funct_treatment("moisture_closest")
num_funct_total("temp_indoor_max")
num_funct_treatment("temp_indoor_max")
num_funct_total("rh_indoor_max")
num_funct_treatment("rh_indoor_max")
num_funct_total("mean_temp")
num_funct_treatment("mean_temp")
num_funct_total("amb_pm_24hr")
num_funct_treatment("amb_pm_24hr")
num_funct_total("sums_events_sampling_period_2.5")
num_funct_treatment("sums_events_sampling_period_2.5")
num_funct_total("sums_events_sampling_period_5")
num_funct_treatment("sums_events_sampling_period_5")
num_funct_total("sums_events_sampling_period_7.5")
num_funct_treatment("sums_events_sampling_period_7.5")

at_home_pm_summary <- health_exposure_data_sampling_day %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>%
  rename(treatment = newtx) %>% 
  arrange(area, home, winter_id, sampling_day) %>% 
  filter(winter_id == 1) %>% 
  distinct(area, child_id_num, .keep_all = TRUE) %>% 
  mutate(new_var = pm_at_home_sampling_period) %>% 
  filter(!is.na(new_var)) %>%
  group_by(treatment) %>% 
  summarize(mean_var = mean(new_var), sd_var = sd(new_var), n_var = n(), 
            min_var = min(new_var), med_var = median(new_var), max_var = max(new_var))
at_home_pm_summary
```

@@

# amb data summary
# amb_pm_lrti, amb_temp_lrti, amb_dewp_lrti, amb_pressure_lrti, amb_wind_lrti
ambient_summary <- analysis_data %>% 
  mutate(new_var = amb_pm_lrti) %>% 
  #group_by(treatment) %>% 
  filter(!is.na(new_var)) %>% 
  #mutate(new_var = (as.numeric(new_var)-32)*5/9) %>% 
  summarize(mean_amb = mean(new_var, na.rm = TRUE),
            sd_amb = sd(new_var, na.rm = TRUE),
            n_amb = n(),
            min_amb = min(new_var, na.rm = TRUE),
            median_amb = median(new_var, na.rm = TRUE),
            max_amb = max(new_var, na.rm = TRUE))
ambient_summary

@@

events_di_plot <- analysis_data %>% 
  ggplot() +
    geom_boxplot(aes(lrti_events_di, person_time_at_risk)) +
    facet_wrap(~treatment)
events_di_plot
```

```{r}
# Check for overdispersion in data using likelihood ratio test
library(pscl)
library(MASS)

nb_results <- glm.nb(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                     data = analysis_data)

poisson_results <- glm(lrti_events ~ treatment + offset(log_person_time_at_risk), 
                       family = "poisson", data = analysis_data)

odTest(nb_results)

@@

inf_results <- influence(nb_results)$hat
inf_results

influence(nb_results)$coef
```

### 2020-10-19 stove_grade_usage_cleaning.Rmd
analysis updates

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 23 June 2020"
output: pdf_document
---

@@

>>>>

---
title: "KidsAIR wood stove grades and usage"
author: "Ethan Walker"
date: "Started 13 Jan 2020, Updated 18 Oct 2020"
output: pdf_document
---

@@

#####################################

# Load, format, and KW monitor data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

##### WMT #####
# Load data
kw_initial_wmt <- read_csv(paste0(file_path, "Input/WMT/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_wmt <- kw_initial_wmt %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "WMT") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_wmt, paste0(file_path, "Output/kw_wmt.rds"))

##### NN #####
# Load data
kw_initial_nn <- read_csv(paste0(file_path, "Input/NN/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_nn <- kw_initial_nn %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "NN") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>% 
  filter(!is.na(kw_hours)) %>%  
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_nn, paste0(file_path, "Output/kw_nn.rds"))

##### AK #####
# Load data
kw_initial_ak <- read_csv(paste0(file_path, "Input/AK/filtretechange.csv"), 
                                  na = c("NULL", "", "...", -99, -999, -9999))

kw_ak <- kw_initial_ak %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         kw_date = ymd(ChangeDate),
         kw_hours = as.numeric(KWHours),
         area = "AK") %>% 
  select(home_winter_id, area, kw_date, kw_hours) %>%  
  filter(!is.na(kw_hours)) %>% 
  ungroup() %>% 
  arrange(home_winter_id, kw_date)

# Save as RDS
write_rds(kw_ak, paste0(file_path, "Output/kw_ak.rds"))
```

# Join KW hour data from different areas
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kw_wmt <- read_rds(paste0(file_path, "Output/kw_wmt.rds"))
kw_nn <- read_rds(paste0(file_path, "Output/kw_nn.rds"))
kw_ak <- read_rds(paste0(file_path, "Output/kw_ak.rds"))

kids_kw <- rbind(kw_wmt, kw_nn, kw_ak) %>% 
  ungroup() %>% 
  arrange(area, home_winter_id, kw_date)

write_rds(kids_kw, paste0(file_path, "Output/kids_kw.rds"))
```

# Clean up all KW hour data
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kw <- read_rds(paste0(file_path, "Output/kids_kw.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds"))
treatments_masked <- read_csv(paste0(file_path, "Input/KidsAIRrnd_masked.csv"), na = "NULL") %>% 
  mutate(home = gsub(" ", "", x = home),
         home = gsub("_", "", x = home))

kids_kw_clean <- kids_kw %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%  
  filter(!is.na(home)) %>% 
  arrange(area, home_winter_id, kw_date) %>% 
  group_by(area, home_winter_id) %>% 
  mutate(kw_hours_total = sum(kw_hours))

summary(kids_kw_clean)

kw_summary <- kids_kw_clean %>% 
  select(-treatment_assigned, -treatment_actual) %>% 
  left_join(treatments_masked, by = "home") %>% 
  distinct(area, home, kw_hours_total) %>% 
  ungroup() %>% 
  #group_by(newtx) %>% 
  summarize(mean(kw_hours_total),
            sd(kw_hours_total),
            n(),
            min(kw_hours_total),
            median(kw_hours_total),
            max(kw_hours_total))
kw_summary

write_rds(kids_kw_clean, paste0(file_path, "Output/kids_kw_clean.rds"))
```
